### 2019-06-09 manuscript/supplement.tex

\documentclass[oneside,10pt,A4]{scrartcl}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{graphicx}

\usepackage[noabbrev]{cleveref}
\usepackage{siunitx}
\DeclareSIUnit \basepair{bp}
\DeclareSIUnit \byte{B}
\DeclareSIPrefix\mebi{Mi}{10}
\DeclareSIPrefix\gibi{Gi}{10}

\input{biocon-stuff.tex}
\input{docker-stuff.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

\usepackage{todonotes}
\setlength{\marginparwidth}{3cm}

\newcounter{todocounter}
\newcommand{\ff}[1]
{\stepcounter{todocounter}
 \todo[color=blue!40,author=For Frank]{\thetodocounter: #1}
 }
\newcommand{\ak}[1]
{\stepcounter{todocounter}
 \todo[color=green!40,author=Arthur]{\thetodocounter: #1}
 }

%%% Put your definitions there:
\newcommand{\formatprogramnames}[1]{\texttt{#1}}
\newcommand{\ce}{\formatprogramnames{chloroExtractor}}
\newcommand{\oa}{\formatprogramnames{ORG.Asm}}
\newcommand{\fp}{\formatprogramnames{Fast-Plast}}
\newcommand{\ioga}{\formatprogramnames{IOGA}}
\newcommand{\np}{\formatprogramnames{NOVOPlasty}}
\newcommand{\go}{\formatprogramnames{GetOrganelle}}
\newcommand{\cassp}{\formatprogramnames{Chloroplast assembly protocol}}

\newcommand{\zenododataset}{\cite{zenododataset}}
\newcommand{\zenodorepo}{\cite{zenodorepo}}

\newcommand{\genename}[1]{\textit{#1}}

% Niklas table
%\newcommand{\ok}{\textcolor[rgb]{0.7,0.7,0}{\textsc{\bfseries{}OKAY}}}
%\newcommand{\bad}{\textcolor{red}{\textsc{\bfseries{}BAD}}}
%\newcommand{\good}{\textcolor[rgb]{0,0.6,0}{\textsc{\bfseries{}GOOD}}}
\newcommand{\ok}{\textsc{Okay}}
\newcommand{\bad}{\textsc{Bad}}
\newcommand{\good}{\textsc{Good}}

\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{capt-of}

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}
%%% Begin ...
\begin{document}

\title{The landscape of chloroplast genome assembly
tools}
\subtitle{Supplemental material}
\author{\parbox{\textwidth}{\centering{}Jan A Freudenthal, Simon Pfaff, Niklas Terhoeven, Arthur Korte, Markus J Ankenbrand and Frank Förster}}
\date{}
\maketitle
\listoftables

\begin{landscape}
\setlength{\LTcapwidth}{1.5\textwidth}
\begin{longtable}{cccrrcc}
\caption[Summary real data set]{\textbf{Summary real data set.} List of real data sets used during our benchmark study. The tables was sorted by species name. Additionally, the table contains the accession for the Sequencing Read Archive (SRA Acc) and the accession of the used reference (Ref Acc). Moreover, the original number of read pairs (read pairs), the overall number of sequenced nucleotides (base pairs), and the mean read length (read len.) is given. Finally, the last column contains the information, if that data set was used for the check of the consistency.} \\
\hline
Taxon & SRA Acc & Ref Acc & read pairs & base pairs (bp) & read len. (bp) & Consistency check \\
\hline
\endfirsthead
\multicolumn{7}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}}\\[0.5em]
Taxon & SRA Acc & Ref Acc & read pairs & base pairs (bp) & read len. (bp) & Consistency check \\ \hline
\endhead
\hline
\multicolumn{7}{r}{Continued on next page}\\
\endfoot
\endlastfoot
\input{sra_ref.tex}
\end{longtable}

\newpage
{\tiny
\captionof{table}{Docker images used in our benchmark setup.}
\label{tab:dockerimages_suppl}
\centering
\begin{tabular}{ccc}
        \hline
          Tool & Image name and tag & SHA256 Checksum   \\ \hline
          \ce & \dockerce & \dockercesha \\
          \cassp & \dockercassp & \dockercasspsha \\
          \fp & \dockerfp & \dockerfpsha \\
          \go & \dockergo & \dockergosha \\
          \ioga & \dockerioga & \dockeriogasha \\
          \np & \dockernp & \dockernpsha \\
          \oa & \dockeroa & \dockeroasha \\ \hline
      \end{tabular}
}
\end{landscape}

\begin{table}[ht]
\caption{Performance metrics input size 25K}
\label{tab:perform25K}
\centering
\begin{tabular}{rlrrrr}
  \hline
 & program & threads & Peak CPU usage & Peak memeory & Peak disk usage (GB) \\ 
  \hline
1 & CAP &   1 & 113.32 & 0.31 & 0.06 \\ 
  2 & CAP &   2 & 166.60 & 0.26 & 0.05 \\ 
  3 & CAP &   4 & 305.99 & 0.26 & 0.05 \\ 
  4 & CAP &   8 & 569.12 & 0.37 & 0.06 \\ 
  5 & CE &   1 & 111.49 & 3.41 & 0.07 \\ 
  6 & CE &   2 & 199.12 & 3.41 & 0.07 \\ 
  7 & CE &   4 & 343.22 & 3.41 & 0.07 \\ 
  8 & CE &   8 & 509.51 & 3.41 & 0.07 \\ 
  9 & Fast-Plast &   1 & 118.31 & 0.77 & 0.30 \\ 
  10 & Fast-Plast &   2 & 213.57 & 0.77 & 0.31 \\ 
  11 & Fast-Plast &   4 & 379.27 & 0.77 & 0.32 \\ 
  12 & Fast-Plast &   8 & 657.20 & 0.77 & 0.35 \\ 
  13 & GetOrganelle &   1 & 111.39 & 0.38 & 0.05 \\ 
  14 & GetOrganelle &   2 & 170.30 & 0.32 & 0.05 \\ 
  15 & GetOrganelle &   4 & 242.27 & 0.31 & 0.06 \\ 
  16 & GetOrganelle &   8 & 346.98 & 0.31 & 0.07 \\ 
  17 & IOGA &   1 & 118.12 & 18.61 & 6.25 \\ 
  18 & IOGA &   2 & 224.74 & 17.88 & 6.26 \\ 
  19 & IOGA &   4 & 394.98 & 16.62 & 5.60 \\ 
  20 & IOGA &   8 & 733.11 & 21.76 & 11.13 \\ 
  21 & NOVOPlasty &   1 & 78.38 & 0.04 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 86.64 & 0.04 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 86.50 & 0.04 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 87.32 & 0.04 & 0.00 \\ 
  25 & org.ASM &   1 & 112.08 & 0.79 & 0.03 \\ 
  26 & org.ASM &   2 & 115.44 & 0.79 & 0.03 \\ 
  27 & org.ASM &   4 & 105.09 & 0.79 & 0.03 \\ 
  28 & org.ASM &   8 & 105.78 & 0.79 & 0.03 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Performance metrics input size 250K}
\label{tab:perform250K}
\centering
\begin{tabular}{rlrrrr}
  \hline
 & program & threads & Peak CPU usage & Peak memeory & Peak disk usage (GB) \\ 
  \hline
1 & CAP &   1 & 114.38 & 0.90 & 0.65 \\ 
  2 & CAP &   2 & 204.39 & 0.93 & 0.65 \\ 
  3 & CAP &   4 & 337.09 & 0.97 & 0.65 \\ 
  4 & CAP &   8 & 652.75 & 1.08 & 0.65 \\ 
  5 & CE &   1 & 117.57 & 3.96 & 0.63 \\ 
  6 & CE &   2 & 230.35 & 3.96 & 0.63 \\ 
  7 & CE &   4 & 414.25 & 3.96 & 0.63 \\ 
  8 & CE &   8 & 755.76 & 3.97 & 0.63 \\ 
  9 & Fast-Plast &   1 & 118.81 & 1.40 & 0.78 \\ 
  10 & Fast-Plast &   2 & 234.12 & 2.25 & 0.78 \\ 
  11 & Fast-Plast &   4 & 425.89 & 2.32 & 0.79 \\ 
  12 & Fast-Plast &   8 & 845.94 & 2.52 & 0.79 \\ 
  13 & GetOrganelle &   1 & 117.94 & 1.07 & 0.40 \\ 
  14 & GetOrganelle &   2 & 227.05 & 1.13 & 0.40 \\ 
  15 & GetOrganelle &   4 & 403.35 & 1.08 & 0.40 \\ 
  16 & GetOrganelle &   8 & 747.20 & 1.03 & 0.39 \\ 
  17 & IOGA &   1 & 118.62 & 12.03 & 4.01 \\ 
  18 & IOGA &   2 & 234.09 & 9.71 & 4.01 \\ 
  19 & IOGA &   4 & 420.77 & 10.09 & 3.48 \\ 
  20 & IOGA &   8 & 842.57 & 19.21 & 9.45 \\ 
  21 & NOVOPlasty &   1 & 111.97 & 0.32 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 115.07 & 0.32 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 116.42 & 0.31 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 115.86 & 0.31 & 0.00 \\ 
  25 & org.ASM &   1 & 118.10 & 0.93 & 0.06 \\ 
  26 & org.ASM &   2 & 161.76 & 0.93 & 0.06 \\ 
  27 & org.ASM &   4 & 163.01 & 0.93 & 0.06 \\ 
  28 & org.ASM &   8 & 160.74 & 0.93 & 0.06 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Performance metrics input size 2.5M}
\label{tab:perform2.5M  }
\centering
\begin{tabular}{rlrrrr}
  \hline
 & program & threads & Peak CPU usage & Peak memeory & Peak disk usage (GB) \\ 
 \hline
1 & CAP &   1 & 118.80 & 6.83 & 6.59 \\ 
  2 & CAP &   2 & 209.22 & 7.58 & 6.59 \\ 
  3 & CAP &   4 & 357.36 & 7.25 & 6.59 \\ 
  4 & CAP &   8 & 683.71 & 6.99 & 6.59 \\ 
  5 & CE &   1 & 118.78 & 5.34 & 2.01 \\ 
  6 & CE &   2 & 234.82 & 5.34 & 2.01 \\ 
  7 & CE &   4 & 415.74 & 5.28 & 2.09 \\ 
  8 & CE &   8 & 844.95 & 5.34 & 2.01 \\ 
  9 & Fast-Plast &   1 & 119.24 & 6.80 & 5.80 \\ 
  10 & Fast-Plast &   2 & 237.34 & 6.98 & 5.82 \\ 
  11 & Fast-Plast &   4 & 443.27 & 8.02 & 5.69 \\ 
  12 & Fast-Plast &   8 & 872.07 & 8.87 & 5.69 \\ 
  13 & GetOrganelle &   1 & 118.91 & 5.25 & 3.31 \\ 
  14 & GetOrganelle &   2 & 235.18 & 5.26 & 3.31 \\ 
  15 & GetOrganelle &   4 & 420.10 & 5.44 & 3.31 \\ 
  16 & GetOrganelle &   8 & 845.13 & 6.07 & 3.31 \\ 
  17 & IOGA &   1 & 119.46 & 18.64 & 8.83 \\ 
  18 & IOGA &   2 & 237.22 & 21.24 & 10.19 \\ 
  19 & IOGA &   4 & 425.50 & 18.71 & 10.56 \\ 
  20 & IOGA &   8 & 857.29 & 25.23 & 16.51 \\ 
  21 & NOVOPlasty &   1 & 112.67 & 2.69 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 117.77 & 2.58 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 117.40 & 2.74 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 117.28 & 2.74 & 0.00 \\ 
  25 & org.ASM &   1 & 118.73 & 2.63 & 0.37 \\ 
  26 & org.ASM &   2 & 184.53 & 2.63 & 0.37 \\ 
  27 & org.ASM &   4 & 177.03 & 2.63 & 0.37 \\ 
  28 & org.ASM &   8 & 176.25 & 2.63 & 0.37 \\ 
   \hline
\end{tabular}
\end{table}

\end{document}