### 2020-05-19 manuscript/supplement.tex

\documentclass[oneside,10pt,A4]{scrartcl}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage{graphicx}

\usepackage{tabularx}
\usepackage{booktabs}

\usepackage[noabbrev]{cleveref}
\usepackage{siunitx}
\DeclareSIUnit \basepair{bp}
\DeclareSIUnit \byte{B}
\DeclareSIPrefix\mebi{Mi}{10}
\DeclareSIPrefix\gibi{Gi}{10}

\input{biocon-stuff.tex}
\input{docker-stuff.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

\usepackage{todonotes}
\setlength{\marginparwidth}{3cm}

\newcounter{todocounter}
\newcommand{\ff}[1]
{\stepcounter{todocounter}
 \todo[color=blue!40,author=For Frank]{\thetodocounter: #1}
 }
\newcommand{\ak}[1]
{\stepcounter{todocounter}
 \todo[color=green!40,author=Arthur]{\thetodocounter: #1}
 }

%%% Put your definitions there:
\newcommand{\formatprogramnames}[1]{\texttt{#1}}
\newcommand{\ce}{\formatprogramnames{chloroExtractor}}
\newcommand{\oa}{\formatprogramnames{ORG.Asm}}
\newcommand{\fp}{\formatprogramnames{Fast-Plast}}
\newcommand{\ioga}{\formatprogramnames{IOGA}}
\newcommand{\np}{\formatprogramnames{NOVOPlasty}}
\newcommand{\go}{\formatprogramnames{GetOrganelle}}
\newcommand{\cassp}{\formatprogramnames{Chloroplast assembly protocol}}

\newcommand{\zenododataset}{\cite{zenododataset}}
\newcommand{\zenodorepo}{\cite{zenodorepo}}

\newcommand{\genename}[1]{\textit{#1}}

% Niklas table
%\newcommand{\ok}{\textcolor[rgb]{0.7,0.7,0}{\textsc{\bfseries{}OKAY}}}
%\newcommand{\bad}{\textcolor{red}{\textsc{\bfseries{}BAD}}}
%\newcommand{\good}{\textcolor[rgb]{0,0.6,0}{\textsc{\bfseries{}GOOD}}}
\newcommand{\ok}{\textsc{Okay}}
\newcommand{\bad}{\textsc{Bad}}
\newcommand{\good}{\textsc{Good}}

\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{capt-of}

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}
%%% Begin ...
\begin{document}

\title{The landscape of chloroplast genome assembly
tools}
\subtitle{Supplemental material}
\author{\parbox{\textwidth}{\centering{}Jan A Freudenthal, Simon Pfaff, Niklas Terhoeven, Arthur Korte, Markus J Ankenbrand and Frank Förster}}
\date{}
\maketitle
\listoftables

\begin{table}[ht]
\caption{Sequencing Read Archive and Genbank accessions for novel chloroplast genomes}
\label{tab:accessions_suppl}
\centering
\begin{tabular}{cccc}\toprule
Species & SRA-Accession & Assembler & Genbank-Accession \\\midrule
\plant{At} & AAAA & AAA & AAAA \\\bottomrule
\end{tabular}
\end{table}

\begin{landscape}
\setlength{\LTcapwidth}{1.5\textwidth}
\begin{longtable}{cccrrcc}
\caption[Summary real data set]{\textbf{Summary real data set.} List of real data sets used during our benchmark study. The tables was sorted by species name. Additionally, the table contains the accession for the Sequencing Read Archive (SRA Acc) and the accession of the used reference (Ref Acc). Moreover, the original number of read pairs (read pairs), the overall number of sequenced nucleotides (base pairs), and the mean read length (read len.) is given. Finally, the last column contains the information, if that data set was used for the check of the consistency.}\label{tab:real_dataset_details} \\
\toprule
Taxon & SRA Acc & Ref Acc & read pairs & base pairs (bp) & read len. (bp) & Consistency check \\\midrule
\endfirsthead
\multicolumn{7}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}}\\[0.5em]
Taxon & SRA Acc & Ref Acc & read pairs & base pairs (bp) & read len. (bp) & Consistency check \\ \midrule
\endhead
\bottomrule
\multicolumn{7}{r}{Continued on next page}\\
\endfoot
\endlastfoot
\input{sra_ref.tex}
\end{longtable}

\newpage
{\tiny
\captionof{table}{Docker images used in our benchmark setup.}
\label{tab:dockerimages_suppl}
\centering
\begin{tabular}{ccc}
        \toprule
          Tool & Image name and tag & SHA256 Checksum   \\ \midrule
          \ce & \dockerce & \dockercesha \\
          \cassp & \dockercassp & \dockercasspsha \\
          \fp & \dockerfp & \dockerfpsha \\
          \go & \dockergo & \dockergosha \\
          \ioga & \dockerioga & \dockeriogasha \\
          \np & \dockernp & \dockernpsha \\
          \oa & \dockeroa & \dockeroasha \\ \bottomrule
      \end{tabular}
}
\end{landscape}

\begin{table}[ht]
\caption{Performance metrics input size 25K}
\label{tab:perform25K_suppl}
\centering
\begin{tabular}{rlrrrr}
  \toprule
 & program & threads & Peak CPU usage & Peak memory & Peak disk usage (GB) \\ 
  \midrule
1 & CAP &   1 & 113.32 & 0.31 & 0.06 \\ 
  2 & CAP &   2 & 166.60 & 0.26 & 0.05 \\ 
  3 & CAP &   4 & 305.99 & 0.26 & 0.05 \\ 
  4 & CAP &   8 & 569.12 & 0.37 & 0.06 \\ 
  5 & CE &   1 & 111.49 & 3.41 & 0.07 \\ 
  6 & CE &   2 & 199.12 & 3.41 & 0.07 \\ 
  7 & CE &   4 & 343.22 & 3.41 & 0.07 \\ 
  8 & CE &   8 & 509.51 & 3.41 & 0.07 \\ 
  9 & Fast-Plast &   1 & 118.31 & 0.77 & 0.30 \\ 
  10 & Fast-Plast &   2 & 213.57 & 0.77 & 0.31 \\ 
  11 & Fast-Plast &   4 & 379.27 & 0.77 & 0.32 \\ 
  12 & Fast-Plast &   8 & 657.20 & 0.77 & 0.35 \\ 
  13 & GetOrganelle &   1 & 111.39 & 0.38 & 0.05 \\ 
  14 & GetOrganelle &   2 & 170.30 & 0.32 & 0.05 \\ 
  15 & GetOrganelle &   4 & 242.27 & 0.31 & 0.06 \\ 
  16 & GetOrganelle &   8 & 346.98 & 0.31 & 0.07 \\ 
  17 & IOGA &   1 & 118.12 & 18.61 & 6.25 \\ 
  18 & IOGA &   2 & 224.74 & 17.88 & 6.26 \\ 
  19 & IOGA &   4 & 394.98 & 16.62 & 5.60 \\ 
  20 & IOGA &   8 & 733.11 & 21.76 & 11.13 \\ 
  21 & NOVOPlasty &   1 & 78.38 & 0.04 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 86.64 & 0.04 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 86.50 & 0.04 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 87.32 & 0.04 & 0.00 \\ 
  25 & org.ASM &   1 & 112.08 & 0.79 & 0.03 \\ 
  26 & org.ASM &   2 & 115.44 & 0.79 & 0.03 \\ 
  27 & org.ASM &   4 & 105.09 & 0.79 & 0.03 \\ 
  28 & org.ASM &   8 & 105.78 & 0.79 & 0.03 \\ 
   \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Performance metrics input size 250K}
\label{tab:perform250K_suppl}
\centering
\begin{tabular}{rlrrrr}
  \toprule
 & program & threads & Peak CPU usage & Peak memory & Peak disk usage (GB) \\ 
  \midrule
1 & CAP &   1 & 114.38 & 0.90 & 0.65 \\ 
  2 & CAP &   2 & 204.39 & 0.93 & 0.65 \\ 
  3 & CAP &   4 & 337.09 & 0.97 & 0.65 \\ 
  4 & CAP &   8 & 652.75 & 1.08 & 0.65 \\ 
  5 & CE &   1 & 117.57 & 3.96 & 0.63 \\ 
  6 & CE &   2 & 230.35 & 3.96 & 0.63 \\ 
  7 & CE &   4 & 414.25 & 3.96 & 0.63 \\ 
  8 & CE &   8 & 755.76 & 3.97 & 0.63 \\ 
  9 & Fast-Plast &   1 & 118.81 & 1.40 & 0.78 \\ 
  10 & Fast-Plast &   2 & 234.12 & 2.25 & 0.78 \\ 
  11 & Fast-Plast &   4 & 425.89 & 2.32 & 0.79 \\ 
  12 & Fast-Plast &   8 & 845.94 & 2.52 & 0.79 \\ 
  13 & GetOrganelle &   1 & 117.94 & 1.07 & 0.40 \\ 
  14 & GetOrganelle &   2 & 227.05 & 1.13 & 0.40 \\ 
  15 & GetOrganelle &   4 & 403.35 & 1.08 & 0.40 \\ 
  16 & GetOrganelle &   8 & 747.20 & 1.03 & 0.39 \\ 
  17 & IOGA &   1 & 118.62 & 12.03 & 4.01 \\ 
  18 & IOGA &   2 & 234.09 & 9.71 & 4.01 \\ 
  19 & IOGA &   4 & 420.77 & 10.09 & 3.48 \\ 
  20 & IOGA &   8 & 842.57 & 19.21 & 9.45 \\ 
  21 & NOVOPlasty &   1 & 111.97 & 0.32 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 115.07 & 0.32 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 116.42 & 0.31 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 115.86 & 0.31 & 0.00 \\ 
  25 & org.ASM &   1 & 118.10 & 0.93 & 0.06 \\ 
  26 & org.ASM &   2 & 161.76 & 0.93 & 0.06 \\ 
  27 & org.ASM &   4 & 163.01 & 0.93 & 0.06 \\ 
  28 & org.ASM &   8 & 160.74 & 0.93 & 0.06 \\ 
   \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{Performance metrics input size 2.5M}
\label{tab:perform2.5M_suppl}
\centering
\begin{tabular}{rlrrrr}
  \toprule
 & program & threads & Peak CPU usage & Peak memory & Peak disk usage (GB) \\ 
 \midrule
1 & CAP &   1 & 118.80 & 6.83 & 6.59 \\ 
  2 & CAP &   2 & 209.22 & 7.58 & 6.59 \\ 
  3 & CAP &   4 & 357.36 & 7.25 & 6.59 \\ 
  4 & CAP &   8 & 683.71 & 6.99 & 6.59 \\ 
  5 & CE &   1 & 118.78 & 5.34 & 2.01 \\ 
  6 & CE &   2 & 234.82 & 5.34 & 2.01 \\ 
  7 & CE &   4 & 415.74 & 5.28 & 2.09 \\ 
  8 & CE &   8 & 844.95 & 5.34 & 2.01 \\ 
  9 & Fast-Plast &   1 & 119.24 & 6.80 & 5.80 \\ 
  10 & Fast-Plast &   2 & 237.34 & 6.98 & 5.82 \\ 
  11 & Fast-Plast &   4 & 443.27 & 8.02 & 5.69 \\ 
  12 & Fast-Plast &   8 & 872.07 & 8.87 & 5.69 \\ 
  13 & GetOrganelle &   1 & 118.91 & 5.25 & 3.31 \\ 
  14 & GetOrganelle &   2 & 235.18 & 5.26 & 3.31 \\ 
  15 & GetOrganelle &   4 & 420.10 & 5.44 & 3.31 \\ 
  16 & GetOrganelle &   8 & 845.13 & 6.07 & 3.31 \\ 
  17 & IOGA &   1 & 119.46 & 18.64 & 8.83 \\ 
  18 & IOGA &   2 & 237.22 & 21.24 & 10.19 \\ 
  19 & IOGA &   4 & 425.50 & 18.71 & 10.56 \\ 
  20 & IOGA &   8 & 857.29 & 25.23 & 16.51 \\ 
  21 & NOVOPlasty &   1 & 112.67 & 2.69 & 0.00 \\ 
  22 & NOVOPlasty &   2 & 117.77 & 2.58 & 0.00 \\ 
  23 & NOVOPlasty &   4 & 117.40 & 2.74 & 0.00 \\ 
  24 & NOVOPlasty &   8 & 117.28 & 2.74 & 0.00 \\ 
  25 & org.ASM &   1 & 118.73 & 2.63 & 0.37 \\ 
  26 & org.ASM &   2 & 184.53 & 2.63 & 0.37 \\ 
  27 & org.ASM &   4 & 177.03 & 2.63 & 0.37 \\ 
  28 & org.ASM &   8 & 176.25 & 2.63 & 0.37 \\ 
   \bottomrule
\end{tabular}
\end{table}

\begin{landscape}
\begin{longtable}{lcccp{7cm}c}
\caption[Novel data set]{\textbf{Novel data set.}}\label{tab:novels_supp} \\
  \toprule
 SRA Accession & Depth & TaxID & Rank & Related References with Depth & Status \\
 \midrule
\endfirsthead
\multicolumn{6}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}}\\[0.5em]
 SRA Accession & Depth & TaxID & Rank & Related References with Depth & Status \\
 \midrule
\endhead
\bottomrule
\multicolumn{6}{r}{Continued on next page}\\
\endfoot
\endlastfoot
DRR053713 & 3 & 71243 & order & NC\_033350(3) & success \\
DRR057122 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
DRR106816 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
ERR1917165 & 5 & 4118 & family & NC\_009808(3) NC\_026703(3) NC\_031159(3) NC\_034670(3) & success \\
ERR2003066 & 3 & 721789 & tribe & NC\_018767(3) NC\_019601(3) NC\_019602(3) NC\_034347(3) NC\_015206(4) NC\_018766(4) & success \\
ERR2744401 & 5 & 156152 & family & NC\_028519(3) NC\_028520(3) NC\_031345(3) NC\_034688(3) NC\_031153(4) NC\_031344(4) & success \\
ERR2744402 & 4 & 156152 & family & NC\_028519(3) NC\_028520(3) NC\_031345(3) NC\_034688(3) NC\_031153(4) NC\_031344(4) & success \\
ERR3089127 & 4 & 102810 & tribe & NC\_020092(3) NC\_020320(3) NC\_020607(3) NC\_025910(3) NC\_030785(3) NC\_031399(3) NC\_031400(3) NC\_034683(3) NC\_034851(3) & success \\
SRR2154064 & 4 & 4210 & family & NC\_027835(4) NC\_034648(4) NC\_007578(5) NC\_007977(5) NC\_010601(5) NC\_013553(5) NC\_015543(5) NC\_015621(5) NC\_020092(5) NC\_020320(5) NC\_020607(5) NC\_023107(5) NC\_023108(5) NC\_023109(5) NC\_023110(5) NC\_023111(5) NC\_023112(5) NC\_023113(5) NC\_023114(5) NC\_023833(5) NC\_024286(5) NC\_025910(5) NC\_027113(5) NC\_027434(5) NC\_028005(5) NC\_028006(5) NC\_028027(5) NC\_029465(5) NC\_030173(5) NC\_030275(5) NC\_030772(5) NC\_030773(5) NC\_030783(5) NC\_030785(5) NC\_031395(5) NC\_031396(5) NC\_031399(5) NC\_031400(5) NC\_031815(5) NC\_031816(5) NC\_031833(5) NC\_031853(5) NC\_031898(5) NC\_032056(5) NC\_032057(5) NC\_034320(5) NC\_034321(5) NC\_034322(5) NC\_034323(5) NC\_034324(5) NC\_034325(5) NC\_034326(5) NC\_034327(5) NC\_034328(5) NC\_034683(5) NC\_034810(5) NC\_034811(5) NC\_034812(5) NC\_034813(5) NC\_034814(5) NC\_034815(5) NC\_034816(5) NC\_034817(5) NC\_034818(5) NC\_034819(5) NC\_034820(5) NC\_034821(5) NC\_034822(5) NC\_034823(5) NC\_034824(5) NC\_034825(5) NC\_034827(5) NC\_034828(5) NC\_034829(5) NC\_034830(5) NC\_034831(5) NC\_034832(5) NC\_034847(5) NC\_034848(5) NC\_034849(5) NC\_034850(5) NC\_034851(5) NC\_034852(5) NC\_034853(5) NC\_034854(5) NC\_034856(5) NC\_034857(5) NC\_034858(5) NC\_034859(5) NC\_034860(5) NC\_034861(5) NC\_034862(5) NC\_034864(5) NC\_034865(5) NC\_034866(5) NC\_034867(5) NC\_034868(5) NC\_034869(5) NC\_034870(5) NC\_034871(5) NC\_034872(5) NC\_034873(5) NC\_034874(5) NC\_034875(5) NC\_034876(5) NC\_034877(5) NC\_034878(5) NC\_034879(5) NC\_034880(5) NC\_034881(5) NC\_034882(5) NC\_034883(5) NC\_034884(5) NC\_034885(5) NC\_034886(5) NC\_034887(5) NC\_034888(5) NC\_034889(5) NC\_034891(5) NC\_034892(5) NC\_034893(5) NC\_034894(5) NC\_034895(5) NC\_034896(5) NC\_034897(5) NC\_034898(5) NC\_034899(5) NC\_034900(5) NC\_034901(5) NC\_034902(5) NC\_034903(5) NC\_034996(5) & success \\
SRR2176133 & 3 & 40550 & order & NC\_026786(3) NC\_027659(3) & success \\
SRR2401791 & 4 & 91888 & no rank & NC\_016433(4) NC\_021449(4) NC\_023463(4) NC\_025652(4) NC\_025653(4) NC\_029719(4) NC\_030212(4) NC\_031435(4) NC\_031436(4) NC\_008407(5) NC\_009808(5) NC\_013707(5) NC\_020098(5) NC\_021111(5) NC\_022859(5) NC\_023102(5) NC\_023115(5) NC\_023131(5) NC\_023132(5) NC\_023464(5) NC\_023465(5) NC\_024845(5) NC\_025641(5) NC\_025642(5) NC\_025651(5) NC\_025787(5) NC\_026202(5) NC\_026291(5) NC\_026703(5) NC\_027262(5) NC\_027838(5) NC\_027955(5) NC\_028519(5) NC\_028520(5) NC\_028533(5) NC\_029700(5) NC\_031159(5) NC\_031345(5) NC\_031437(5) NC\_031441(5) NC\_031442(5) NC\_031443(5) NC\_031444(5) NC\_031445(5) NC\_033534(5) NC\_034308(5) NC\_034309(5) NC\_034310(5) NC\_034311(5) NC\_034312(5) NC\_034670(5) NC\_034688(5) NC\_034691(5) NC\_034693(5) NC\_034694(5) NC\_034695(5) NC\_035000(5) NC\_001879(6) NC\_004561(6) NC\_007500(6) NC\_007602(6) NC\_007943(6) NC\_008096(6) NC\_015401(6) NC\_015604(6) NC\_015608(6) NC\_015623(6) NC\_016068(6) NC\_018117(6) NC\_018552(6) NC\_022451(6) NC\_024261(6) NC\_024292(6) NC\_025657(6) NC\_026551(6) NC\_026563(6) NC\_026567(6) NC\_026570(6) NC\_026574(6) NC\_026694(6) NC\_026726(6) NC\_026906(6) NC\_027099(6) NC\_027177(6) NC\_027441(6) NC\_027442(6) NC\_028007(6) NC\_028009(6) NC\_028069(6) NC\_028070(6) NC\_028614(6) NC\_029370(6) NC\_029746(6) NC\_029817(6) NC\_029818(6) NC\_029819(6) NC\_029820(6) NC\_029821(6) NC\_029822(6) NC\_029823(6) NC\_029824(6) NC\_029825(6) NC\_029833(6) NC\_030044(6) NC\_030056(6) NC\_030167(6) NC\_030168(6) NC\_030171(6) NC\_030177(6) NC\_030178(6) NC\_030185(6) NC\_030207(6) NC\_030282(6) NC\_030319(6) NC\_030543(6) NC\_030755(6) NC\_030756(6) NC\_030757(6) NC\_031153(6) NC\_031155(6) NC\_031344(6) NC\_031433(6) NC\_031434(6) NC\_031805(6) NC\_031874(6) NC\_032054(6) NC\_032724(6) NC\_033354(6) NC\_033524(6) NC\_033525(6) NC\_033526(6) NC\_033527(6) NC\_034698(6) NC\_034951(6) NC\_035143(6) AC\_000188(7) NC\_007898(7) NC\_016468(7) NC\_020431(7) NC\_021423(7) NC\_022431(7) NC\_022432(7) NC\_025655(7) NC\_025656(7) NC\_025658(7) NC\_026876(7) NC\_026877(7) NC\_026878(7) NC\_026879(7) NC\_026880(7) NC\_026881(7) NC\_026882(7) NC\_027259(7) NC\_029459(7) NC\_029460(7) NC\_031852(7) NC\_034660(7) NC\_035233(7) NC\_008535(8) NC\_030053(8) & success \\
SRR2531270 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR2531276 & 4 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR2531278 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR2531280 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR2531284 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR2531289 & 3 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR4124062 & 7 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR4124063 & 6 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR4124078 & 6 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR4124079 & 7 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR4457826 & 4 & 102809 & tribe & NC\_027434(3) NC\_034810(3) NC\_034811(3) NC\_034812(3) NC\_034813(3) NC\_034814(3) NC\_034815(3) NC\_034816(3) NC\_034817(3) NC\_034818(3) NC\_034819(3) NC\_034820(3) NC\_034821(3) NC\_034822(3) NC\_034823(3) NC\_034824(3) NC\_034825(3) NC\_034827(3) NC\_034828(3) NC\_034829(3) NC\_034830(3) NC\_034831(3) NC\_034832(3) NC\_034847(3) NC\_034848(3) NC\_034849(3) NC\_034850(3) NC\_034852(3) NC\_034853(3) NC\_034854(3) NC\_034856(3) NC\_034857(3) NC\_034858(3) NC\_034859(3) NC\_034860(3) NC\_034861(3) NC\_034862(3) NC\_034864(3) NC\_034865(3) NC\_034866(3) NC\_034867(3) NC\_034868(3) NC\_034869(3) NC\_034870(3) NC\_034871(3) NC\_034872(3) NC\_034873(3) NC\_034874(3) NC\_034875(3) NC\_034876(3) NC\_034877(3) NC\_034878(3) NC\_034879(3) NC\_034880(3) NC\_034881(3) NC\_034882(3) NC\_034883(3) NC\_034884(3) NC\_034885(3) NC\_034886(3) NC\_034887(3) NC\_034888(3) NC\_034889(3) NC\_034891(3) NC\_034892(3) NC\_034893(3) NC\_034894(3) NC\_034895(3) NC\_034896(3) NC\_034897(3) NC\_034898(3) NC\_034899(3) NC\_034900(3) NC\_034901(3) NC\_034902(3) NC\_034903(3) NC\_034996(3) & success \\
SRR4457835 & 4 & 102809 & tribe & NC\_027434(3) NC\_034810(3) NC\_034811(3) NC\_034812(3) NC\_034813(3) NC\_034814(3) NC\_034815(3) NC\_034816(3) NC\_034817(3) NC\_034818(3) NC\_034819(3) NC\_034820(3) NC\_034821(3) NC\_034822(3) NC\_034823(3) NC\_034824(3) NC\_034825(3) NC\_034827(3) NC\_034828(3) NC\_034829(3) NC\_034830(3) NC\_034831(3) NC\_034832(3) NC\_034847(3) NC\_034848(3) NC\_034849(3) NC\_034850(3) NC\_034852(3) NC\_034853(3) NC\_034854(3) NC\_034856(3) NC\_034857(3) NC\_034858(3) NC\_034859(3) NC\_034860(3) NC\_034861(3) NC\_034862(3) NC\_034864(3) NC\_034865(3) NC\_034866(3) NC\_034867(3) NC\_034868(3) NC\_034869(3) NC\_034870(3) NC\_034871(3) NC\_034872(3) NC\_034873(3) NC\_034874(3) NC\_034875(3) NC\_034876(3) NC\_034877(3) NC\_034878(3) NC\_034879(3) NC\_034880(3) NC\_034881(3) NC\_034882(3) NC\_034883(3) NC\_034884(3) NC\_034885(3) NC\_034886(3) NC\_034887(3) NC\_034888(3) NC\_034889(3) NC\_034891(3) NC\_034892(3) NC\_034893(3) NC\_034894(3) NC\_034895(3) NC\_034896(3) NC\_034897(3) NC\_034898(3) NC\_034899(3) NC\_034900(3) NC\_034901(3) NC\_034902(3) NC\_034903(3) NC\_034996(3) & success \\
SRR5195283 & 3 & 4055 & order & NC\_024292(5) NC\_025657(5) NC\_027441(5) NC\_027442(5) NC\_028009(5) NC\_028614(5) NC\_030319(5) NC\_031155(5) NC\_033354(5) NC\_034698(5) NC\_021423(6) NC\_022431(6) NC\_022432(6) NC\_025655(6) NC\_025656(6) NC\_025658(6) NC\_029459(6) NC\_029460(6) NC\_008535(7) NC\_030053(7) & success \\
SRR5409224 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR5409225 & 4 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR5581778 & 4 & 3650 & family & NC\_007144(3) NC\_023544(3) NC\_029484(3) NC\_031834(3) NC\_032008(3) NC\_033899(3) NC\_015983(4) & success \\
SRR5586278 & 5 & 1521262 & subclass & NC\_012818(4) NC\_022136(4) NC\_022137(4) NC\_024157(4) NC\_024158(4) NC\_004766(6) NC\_014348(6) NC\_014592(6) NC\_028542(6) NC\_028543(6) NC\_028705(6) & success \\
SRR5880481 & 3 & 41945 & order & NC\_020019(3) NC\_021121(3) NC\_022264(3) NC\_022459(3) NC\_022460(3) NC\_022461(3) NC\_022462(3) NC\_022463(3) NC\_023084(3) NC\_024541(3) NC\_024543(3) NC\_024659(3) NC\_024660(3) NC\_024661(3) NC\_024662(3) NC\_024663(3) NC\_026197(3) NC\_026690(3) NC\_026691(3) NC\_030180(3) NC\_030539(3) NC\_030609(3) NC\_030784(3) NC\_030786(3) NC\_030787(3) NC\_030789(3) NC\_031186(3) NC\_031187(3) NC\_031428(3) NC\_031892(3) NC\_033502(3) NC\_034331(3) NC\_034371(3) NC\_034640(3) NC\_034641(3) NC\_034677(3) NC\_034678(3) NC\_034913(3) NC\_034914(3) NC\_034915(3) NC\_035418(3) NC\_033501(4) NC\_019616(5) NC\_029704(5) & success \\
SRR6331521 & 3 & 4149 & family & NC\_026202(3) NC\_031437(3) & success \\
SRR6425047 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR6425048 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR6425049 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR6425628 & 4 & 167487 & subfamily & NC\_024292(3) NC\_033354(3) NC\_021423(4) & success \\
SRR6425646 & 4 & 167487 & subfamily & NC\_024292(3) NC\_033354(3) NC\_021423(4) & success \\
SRR6425647 & 4 & 167487 & subfamily & NC\_024292(3) NC\_033354(3) NC\_021423(4) & success \\
SRR6425648 & 4 & 167487 & subfamily & NC\_024292(3) NC\_033354(3) NC\_021423(4) & success \\
SRR6425651 & 4 & 167483 & subfamily & NC\_025657(3) NC\_025655(4) NC\_025656(4) NC\_025658(4) & success \\
SRR6513658 & 3 & 3814 & subfamily & NC\_009259(6) NC\_013843(6) NC\_016708(6) NC\_018051(6) NC\_021091(6) NC\_023090(6) NC\_025909(6) NC\_027677(6) NC\_029406(6) NC\_031429(6) NC\_034742(6) NC\_034743(6) NC\_034774(6) NC\_002694(7) NC\_003119(7) NC\_007942(7) NC\_011163(7) NC\_011828(7) NC\_014057(7) NC\_014063(7) NC\_021636(7) NC\_021645(7) NC\_021646(7) NC\_021647(7) NC\_021648(7) NC\_021649(7) NC\_021650(7) NC\_022868(7) NC\_024034(7) NC\_024035(7) NC\_024036(7) NC\_024038(7) NC\_024166(7) NC\_025743(7) NC\_025744(7) NC\_025745(7) NC\_027073(7) NC\_027074(7) NC\_027075(7) NC\_027076(7) NC\_027077(7) NC\_027078(7) NC\_027079(7) NC\_027080(7) NC\_027148(7) NC\_027149(7) NC\_027150(7) NC\_027151(7) NC\_027152(7) NC\_027153(7) NC\_027154(7) NC\_027155(7) NC\_029828(7) NC\_030329(7) NC\_032066(7) NC\_032691(7) NC\_034229(7) NC\_035228(7) NC\_035229(7) NC\_028171(8) & success \\
SRR6821984 & 3 & 3700 & family & NC\_000932(3) NC\_009265(3) NC\_009266(3) NC\_009267(3) NC\_009268(3) NC\_009269(3) NC\_009270(3) NC\_009271(3) NC\_009272(3) NC\_009273(3) NC\_009274(3) NC\_009275(3) NC\_016734(3) NC\_018565(3) NC\_021102(3) NC\_023367(3) NC\_024469(3) NC\_026445(3) NC\_026446(3) NC\_027693(3) NC\_028170(3) NC\_028272(3) NC\_028415(3) NC\_028517(3) NC\_028726(3) NC\_028727(3) NC\_028728(3) NC\_029253(3) NC\_029254(3) NC\_029331(3) NC\_029332(3) NC\_029333(3) NC\_029334(3) NC\_029335(3) NC\_029336(3) NC\_029337(3) NC\_029378(3) NC\_029379(3) NC\_030346(3) NC\_030347(3) NC\_030348(3) NC\_030349(3) NC\_030350(3) NC\_030351(3) NC\_030450(3) NC\_030515(3) NC\_030516(3) NC\_030775(3) NC\_033499(3) NC\_033500(3) NC\_034287(3) NC\_034299(3) NC\_034357(3) NC\_034359(3) NC\_034360(3) NC\_034361(3) NC\_034362(3) NC\_034363(3) NC\_034365(3) NC\_034366(3) NC\_034367(3) NC\_035303(3) NC\_034286(4) NC\_034379(4) & success \\
SRR6940055 & 3 & 4136 & family & NC\_020098(3) NC\_023102(3) NC\_026291(3) NC\_027262(3) NC\_028533(3) NC\_034693(3) NC\_029370(4) NC\_029817(4) NC\_029818(4) NC\_029819(4) NC\_029820(4) NC\_029821(4) NC\_029822(4) NC\_029823(4) NC\_029824(4) NC\_029825(4) NC\_030755(4) NC\_030756(4) NC\_030757(4) NC\_031433(4) NC\_031434(4) NC\_031874(4) NC\_032054(4) NC\_035143(4) NC\_020431(5) NC\_027259(5) NC\_035233(5) & success \\
SRR6940081 & 3 & 4136 & family & NC\_020098(3) NC\_023102(3) NC\_026291(3) NC\_027262(3) NC\_028533(3) NC\_034693(3) NC\_029370(4) NC\_029817(4) NC\_029818(4) NC\_029819(4) NC\_029820(4) NC\_029821(4) NC\_029822(4) NC\_029823(4) NC\_029824(4) NC\_029825(4) NC\_030755(4) NC\_030756(4) NC\_030757(4) NC\_031433(4) NC\_031434(4) NC\_031874(4) NC\_032054(4) NC\_035143(4) NC\_020431(5) NC\_027259(5) NC\_035233(5) & success \\
SRR7223704 & 3 & 3700 & family & NC\_000932(3) NC\_009265(3) NC\_009266(3) NC\_009267(3) NC\_009268(3) NC\_009269(3) NC\_009270(3) NC\_009271(3) NC\_009272(3) NC\_009273(3) NC\_009274(3) NC\_009275(3) NC\_016734(3) NC\_018565(3) NC\_021102(3) NC\_023367(3) NC\_024469(3) NC\_026445(3) NC\_026446(3) NC\_027693(3) NC\_028170(3) NC\_028272(3) NC\_028415(3) NC\_028517(3) NC\_028726(3) NC\_028727(3) NC\_028728(3) NC\_029253(3) NC\_029254(3) NC\_029331(3) NC\_029332(3) NC\_029333(3) NC\_029334(3) NC\_029335(3) NC\_029336(3) NC\_029337(3) NC\_029378(3) NC\_029379(3) NC\_030346(3) NC\_030347(3) NC\_030348(3) NC\_030349(3) NC\_030350(3) NC\_030351(3) NC\_030450(3) NC\_030515(3) NC\_030516(3) NC\_030775(3) NC\_033499(3) NC\_033500(3) NC\_034287(3) NC\_034299(3) NC\_034357(3) NC\_034359(3) NC\_034360(3) NC\_034361(3) NC\_034362(3) NC\_034363(3) NC\_034365(3) NC\_034366(3) NC\_034367(3) NC\_035303(3) NC\_034286(4) NC\_034379(4) & success \\
SRR7223710 & 3 & 3700 & family & NC\_000932(3) NC\_009265(3) NC\_009266(3) NC\_009267(3) NC\_009268(3) NC\_009269(3) NC\_009270(3) NC\_009271(3) NC\_009272(3) NC\_009273(3) NC\_009274(3) NC\_009275(3) NC\_016734(3) NC\_018565(3) NC\_021102(3) NC\_023367(3) NC\_024469(3) NC\_026445(3) NC\_026446(3) NC\_027693(3) NC\_028170(3) NC\_028272(3) NC\_028415(3) NC\_028517(3) NC\_028726(3) NC\_028727(3) NC\_028728(3) NC\_029253(3) NC\_029254(3) NC\_029331(3) NC\_029332(3) NC\_029333(3) NC\_029334(3) NC\_029335(3) NC\_029336(3) NC\_029337(3) NC\_029378(3) NC\_029379(3) NC\_030346(3) NC\_030347(3) NC\_030348(3) NC\_030349(3) NC\_030350(3) NC\_030351(3) NC\_030450(3) NC\_030515(3) NC\_030516(3) NC\_030775(3) NC\_033499(3) NC\_033500(3) NC\_034287(3) NC\_034299(3) NC\_034357(3) NC\_034359(3) NC\_034360(3) NC\_034361(3) NC\_034362(3) NC\_034363(3) NC\_034365(3) NC\_034366(3) NC\_034367(3) NC\_035303(3) NC\_034286(4) NC\_034379(4) & success \\
SRR8082590 & 3 & 3646 & order & NC\_024060(3) NC\_024061(3) NC\_024062(3) NC\_024063(3) NC\_024064(3) NC\_024065(3) NC\_024066(3) NC\_024067(3) NC\_026986(3) NC\_030517(3) NC\_030534(3) NC\_030544(3) NC\_030545(3) NC\_030546(3) NC\_030547(3) NC\_030548(3) NC\_030549(3) NC\_030550(3) NC\_030551(3) NC\_030552(3) NC\_030553(3) NC\_030554(3) NC\_030555(3) NC\_030556(3) NC\_030557(3) NC\_030558(3) NC\_030559(3) NC\_030560(3) NC\_030561(3) NC\_030562(3) NC\_030563(3) NC\_030564(3) NC\_030565(3) NC\_030566(3) NC\_030567(3) NC\_030568(3) NC\_030569(3) NC\_030570(3) NC\_030571(3) NC\_030572(3) NC\_030573(3) NC\_030574(3) NC\_030575(3) NC\_030576(3) NC\_030577(3) NC\_030578(3) NC\_030579(3) NC\_030580(3) NC\_030581(3) NC\_030582(3) NC\_030583(3) NC\_030601(3) NC\_034285(3) NC\_008235(4) NC\_009143(4) NC\_024681(4) NC\_024734(4) NC\_024735(4) NC\_024747(4) NC\_026462(4) NC\_026722(4) NC\_027425(4) NC\_028350(4) NC\_028504(4) NC\_031371(4) NC\_031398(4) NC\_032060(4) NC\_032368(4) NC\_032717(4) NC\_033876(4) NC\_010433(5) NC\_012224(5) NC\_015308(5) NC\_016736(5) NC\_034803(5) NC\_033910(7) & success \\
SRR8481357 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR8481359 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR8784049 & 7 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
SRR8784093 & 5 & 2231393 & no rank & NC\_009259(5) NC\_013843(5) NC\_016708(5) NC\_018051(5) NC\_021091(5) NC\_023090(5) NC\_025909(5) NC\_027677(5) NC\_029406(5) NC\_031429(5) NC\_034742(5) NC\_034743(5) NC\_034774(5) NC\_002694(6) NC\_003119(6) NC\_007942(6) NC\_011163(6) NC\_011828(6) NC\_014057(6) NC\_014063(6) NC\_021636(6) NC\_021645(6) NC\_021646(6) NC\_021647(6) NC\_021648(6) NC\_021649(6) NC\_021650(6) NC\_022868(6) NC\_024034(6) NC\_024035(6) NC\_024036(6) NC\_024038(6) NC\_024166(6) NC\_025743(6) NC\_025744(6) NC\_025745(6) NC\_027073(6) NC\_027074(6) NC\_027075(6) NC\_027076(6) NC\_027077(6) NC\_027078(6) NC\_027079(6) NC\_027080(6) NC\_027148(6) NC\_027149(6) NC\_027150(6) NC\_027151(6) NC\_027152(6) NC\_027153(6) NC\_027154(6) NC\_027155(6) NC\_029828(6) NC\_030329(6) NC\_032066(6) NC\_032691(6) NC\_034229(6) NC\_035228(6) NC\_035229(6) NC\_028171(7) & success \\
\end{longtable}
\end{landscape}
\end{document}